Boosting Higgs Pair Production in the $b\bar{b}b\bar{b}$ Final State with Multivariate Techniques

### 2015-12-15 paper/HH4b.tex
working on the text

\abstract{
  The measurement of Higgs pair production will be a cornerstone of
  the LHC program in the coming years.
  %
  Double Higgs production provides a crucial window upon the mechanism of electroweak
  symmetry breaking and has a unique sensitivity
  to the Higgs trilinear coupling.
  %
  In this work, we study the feasibility of a measurement of
  Higgs pair production in the
  $b\bar{b}b\bar{b}$ final state at the LHC.
  %
  Our analysis is based on a combination of traditional cut-based
  methods with state-of-the-art
  multivariate techniques.
  %
  We account for all relevant
  backgrounds, including the contributions from light and charm
  jet mistags, which are ultimately comparable in size 
  to the irreducible $4b$ QCD background.
  %
  We demonstrate the robustness of our analysis strategy against the
  effects of pile-up.
  %
  For an integrated luminosity of $\mathcal{L}=3$ ab$^{-1}$,
  a signal significance of  $S/\sqrt{B}\simeq 4$ is obtained,
  indicating
  that the $b\bar{b}b\bar{b}$ channel
alone could allow for the observation of double Higgs production
  at the High Luminosity LHC.
}

>>>>

\abstract{
  The measurement of Higgs pair production will be a cornerstone of
  the LHC program in the coming years.
  %
  Double Higgs production provides a crucial window upon the mechanism of electroweak
  symmetry breaking and has a unique sensitivity
  to the Higgs trilinear coupling.
  %
  We study the feasibility of a measurement of
  Higgs pair production in the
  $b\bar{b}b\bar{b}$ final state at the LHC.
  %
  Our analysis is based on a combination of traditional cut-based
  methods with state-of-the-art
  multivariate techniques.
  %
  We account for all relevant
  backgrounds, including the contributions from light and charm
  jet mis-identification, which are ultimately comparable in size 
  to the irreducible $4b$ QCD background.
  %
  We demonstrate the robustness of our analysis strategy against the
  effects of pile-up.
  %
  For an integrated luminosity of $\mathcal{L}=3$ ab$^{-1}$,
  a signal significance of  $S/\sqrt{B}\simeq 4$ is obtained,
  indicating
  that the $b\bar{b}b\bar{b}$ channel
alone could allow for the observation of double Higgs production
  at the High Luminosity LHC.
}

### 2015-12-15 paper/sec-analysis.tex
working on the text

In this section we describe our analysis strategy.
%
First of all we discuss the settings
for jet clustering and the strategy for jet $b$-tagging.
%
Following this we discuss the categorisation of events into different
topologies, and how the different topologies may be prioritised.
%
We motivate our choice of analysis cuts by comparing signal and background
distributions for representative kinematic variables.
%
Finally, we describe the simulation of PU and validate
the PU subtraction strategy that we adopt.

@@

  The $N$-subjettiness variables $\tau_N$ are defined by clustering the constituents
  of a jet with the exclusive $k_t$ algorithm~\cite{Catani:1993hr}
  and requiring that $N$ subjets are found,
  \be
  \tau_N \equiv \frac{1}{d_0} \sum_k p_{T,k}\cdot {\rm min}\lp \delta R_{1k}, \ldots,
  \delta R_{Nk}\rp \, , \qquad d_0\equiv \sum_k p_{T,k}\cdot R \, ,
  \ee
  where $p_{T,k}$ is the $p_T$ of the constituent particle $k$ and $\delta R_{ik}$ the distance from
  subjet $i$ to constituent $k$.
  %
  In this work we use as input to the MVA the ratio of 2-subjettiness to 1-subjettiness
  \be
  \label{eq:tau21}
\tau_{21} \equiv \frac{\tau_2}{\tau_1} \, ,
  \ee
  which provides good discrimination 
  between QCD jets and jets arising from the decay of
  a heavy resonance.

@@

    Large-$R$ jets are $b$-tagged by
    ghost-associating anti-$k_T$ $R=0.3$ (AKT03)
    subjets to the original large-$R$
    jets~\cite{Cacciari:2007fd,Aad:2013gja,
      ATLAS-CONF-2014-004,Aad:2015uka}.
    %
    A large-$R$ jet is considered $b$-tagged if both
    the leading and subleading AK03 subjets, where the ordering
    is done in the subjet $p_T$, are both individually $b$-tagged,
    with the same criteria as the small-$R$ jets.
    %
     Therefore, a large-$R$ jet where the two leading
    subjets have at least one $b$-quark will be tagged
    with probability $f_b^2$.

@@

For the $b$-tagging probability $f_b$, along with
the $b$-mistag probability of light ($f_l$) and charm ($f_c$) jets,
we used the values $f_b=0.8$, $f_l=0.01$
and  $f_c=0.1$.

@@

  An event which
  contains at least two large-$R$ jets, with the two leading jets
 being $b$-tagged.
 %
 Each of these two large-$R$, $b$-tagged, jets are 
 therefore candidates
 to contain the decay products of a Higgs boson.

@@

  An event with exactly one large-$R$, $b$-tagged jet, which
  is assigned to be the leading Higgs candidate.
  %
  In addition, we require at least two small-$R$, $b$-tagged, jets,
  which must be separated with respect to the large-$R$ jet
  by an angular distance of $\Delta R\ge 1.2$.
  %

@@

The separate optimisation of the three categories needs to be performed in
a inclusive analysis, following which we introduce a
 prioritisation whereby events are assigned an exclusive category starting 
 with the category of highest signal significance.

 This has been achieved as follows.
 %
First of all we perform an inclusive analysis, and optimise the
signal significance
$S/\sqrt{B}$ in each of the three categories separately, including
the MVA.
%
We find that the category with highest significance is
the boosted one,
followed by the intermediate and the resolved topologies, the latter two
with similar significance.
%
Therefore, when ascertaining which category an event is to be exclusively placed:
if the event satisfies the boosted requirements, it is assigned to
this category, else we check if it suits the intermediate
requirements.
%
It the event also fails the intermediate category
requirements, we
then check if it pass the resolved selection criteria.
%
The resulting exclusive event samples are then separately processed
through the MVA, allowing a consistent combination
of the significance of the three event categories.

@@

We now motivate the
kinematic cuts applied to the different categories, 
comparing representative kinematic distributions between
signal and background events.
%
Firstly we present results without PU, and then
discuss
the impact of PU
on the description of the various kinematic
distributions.
%
In the following, all
distributions will be normalized to their total integral.

In Fig.~\ref{fig:cutplots1} we show
the $p_T$ distributions
of the
  leading and subleading large-$R$ jets in the boosted category, comparing
  the shapes of the distribution in signal and background events.
  %
  We observe that the background distribution
falls off more rapidly as a function of $p_T$ than the di-Higgs signal.
  %
  On the other hand, the cut in $p_T$ cannot be too strong to avoid
  a substantial degradation of signal selection efficiency,
  specially taking into account the subleading large-$R$ jet.
  %
  This comparison justifies the cut of $p_T \ge 200$ GeV
  for the large-$R$ jets that we impose in the boosted category.

@@

Another selection requirement for the boosted category is that the two
leading AKT03 subjets of the large-$R$ jet must be relatively hard,
in particular they should satisfy $p_T \ge $ 50 GeV.
%
To motivate this cut, in Fig.~\ref{fig:cutplots22}
we show the distribution in $p_T$ of the leading
and subleading AKT03 subjets in the subleading large-$R$ jet in events
corresponding to the boosted category.
%
It is clear from the comparison that the subjet $p_T$ spectrum is
relatively harder in the signal with respect to the background.
%
On the other hand, considering the subleading AKT03 subjet,
this cut in $p_T$
cannot be too harsh, to maintain a high signal selection
efficiency.
%
Therefore,
as for the previous distribution, also here 
a compromise between suppressing backgrounds but keeping a large fraction of
signal events is crucial.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
 \includegraphics[width=0.48\textwidth]{plots/pt_leadSJ_fj2_noPU.pdf}
 \includegraphics[width=0.48\textwidth]{plots/pt_subleadSJ_fj2_noPU.pdf}
 \caption{\small  Same as Fig.~\ref{fig:cutplots1} for the two leading AKT03
   subjets in the subleading Higgs candidate large-$R$ jet.
}
\label{fig:cutplots22}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

Turning to the resolved category, an important aspect to account for
in the corresponding selection
cuts is the fact that the $p_T$ distribution
of the four leading small-$R$ jets of the event can be relatively soft,
specially for the subleading jets.
%
As noted in~\cite{deLima:2014dta}, this is a consequence of the fact
that in general the boost from the Higgs decays is moderate.
Therefore the $p_T$ selection cuts for the small-$R$ jets cannot be too large.
%
In Fig.~\ref{fig:cutplots23}
we show the distribution in $p_T$ of the four leading
small-$R$ jets in signal and background events: we observe that both
distributions peak at $p_T \le 50$ GeV, with the signal distribution
eventually becoming dominant at large $p_T$.
%
The feasibility of triggering on four small-$R$ jets with a relatively
soft $p_T$ distribution is one of the experimental challenges for
exploiting the resolved category in this final state,
and hence the requirement that $p_T \ge 40$ GeV for
the small-$R$ jets.
%
In  Fig.~\ref{fig:cutplots23} we also show the
rapidity distribution of the the small-$R$
jets in the resolved category.
%
As expected, the production
is mostly central, more so in the case
of signal events, since backgrounds are dominated by
QCD $t$-channel exchange, therefore the
selection criteria on the jet rapidity are very efficient.

@@

The invariant mass of the di-Higgs system is another important
kinematic distribution for this process.
%
Thie di-Higgs invariant mass is a direct measure of the boost of the system,
which in  BSM scenarios can be substantially
enhanced, for instance due to
specific $d=6$ EFT operators~\cite{Azatov:2015oxa}.
%
One important advantage of the $b\bar{b}b\bar{b}$ final state for
di-Higgs production is that it significantly increases the reach
in $m_{hh}$ as compared to other channels with smaller branching
ratios,
such as $2b2\gamma$ or $2b2\tau$.
%
In Fig.~\ref{fig:mhh} we show the invariant mass distribution of the
reconstructed Higgs pairs,
comparing the resolved and the boosted categories.

@@

In the resolved case, we see that the distribution
in $m_{hh}$ is rather harder for the signal as compared
to the background,
and therefore one expects that cutting in $m_{hh}$ would help signal
discrimination: we will verify this using the MVA.
%
For the boosted category the overall trend of the $m_{hh}$ distribution
is different because of the jet selection criteria, and the
distribution now peaks at higher values of the invariant mass.
%
In this case, signal and background distributions are not significantly
differentiated.
%
Note that at parton-level the $m_{hh}$ distribution
for signal events has a kinematic
cut-off at $m_{hh}^{\rm min}=250$ GeV, which is smeared due
to 
the parton shower and to detector resolution effects.
%

In Fig.~\ref{fig:pthh} we show the transverse momentum of
the di-Higgs system, $p_T^{hh}$,
for the resolved and boosted categories.
%
Once more we see that the background has a steeper fall-off in $p_T^{hh}$
than the signal, in both categories, therefore this variable
should provide additional discrimination power, motivating it's inclusion
as one of the inputs for the MVA.
%
In our LO simulation the $p_T^{hh}$ distribution is generated
by the parton shower, and improved theoretical
description would require
either the matching with higher-multiplicity
matrix elements~\cite{Maierhofer:2013sha} or
the full NLO calculation~\cite{Frederix:2014hta},
%
Nevertheless, as we will show,
the MVA shows only limited sensitivity to this variable, so its
modelling is not crucial in this specific application.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
    \includegraphics[width=0.48\textwidth]{plots/split12_h1_C1_boost.pdf} 
  \includegraphics[width=0.48\textwidth]{plots/EEC_C2_h0_C1_boost.pdf}
  \includegraphics[width=0.48\textwidth]{plots/tau21_h0_C1_boost.pdf}
  \includegraphics[width=0.48\textwidth]{plots/tau21_h1_C1_boost.pdf}
  \caption{\small Distribution of representative substructure variables
    in the boosted category at the end of the cut-based
    analysis, to be used as input to the MVA.
    %
    From top to bottom and from left to right we show  the
    $k_t$ splitting scale, the energy correlation ratio $C_2^{(\beta)}$
    and the  subjettiness ratio $\tau_{21}$ for the leading
    Higgs.
    %
    In the case of 
    $\tau_{21}$  the distributions for the subleading Higgs are also given.
 }
\label{fig:mva_substructure_1}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

From Fig.~\ref{fig:mva_substructure_1}
we can observe how for these substructure variables the shapes of the signal
and background distributions reflect
the inherent differences in the internal structure of jets
between QCD jets and jets originating from the Higgs decays.
%
Signal and background distributions peak
in rather
different regions. For example, the $k_t$ splitting scale $\sqrt{d_{12}}$
peaks around 80 GeV (40 GeV) for signal (background) events, while
the distribution of the
ECF ratio $C_2^{(\beta)}$ is concentrated at small values
for signal but is much broader for background events.
%
From Fig.~\ref{fig:mva_substructure_1} we also see
the distributions of the subjettiness ratio $\tau_{12}$ are
reasonably similar
for both the leading and the subleading jets.
%
Therefore, while we do not explicitly have cuts in the substructure
variables, their discrimination power between signal and
background events is the motivation to include them
as input to the MVA, as discussed in Sect.~\ref{sec:mva}.

@@

Now we turn to discuss how the description of kinematic
distributions for $hh\to 4b$ signal
and background processes are
modified in the presence of pile-up (PU).
%
To study the impact of PU,
Minimum Bias (MB) events,
including Multiple Parton Interactions (MPI), have been generated
with {\tt Pythia8}, and then
superimposed to the signal
and background samples described in Sect.~\ref{mcgeneration}.
%
We have explored two scenarios,
one with a mean number of
PU vertices per bunch crossing of $\la n_{\rm PU}\ra=80$, and another
with $\la n_{\rm PU}\ra=150$.
%
While the latter is closer to the HL-LHC forecasts,
we adopt $\la n_{\rm PU}\ra=80$ as our baseline,
given that it
is beyond the scope of this paper to fully optimize
the PU subtraction.

@@

In Fig.~\ref{fig:mHH_PU}
we compare the transverse momentum of the leading Higgs
candidate, $p_t^{h_1}$ and the invariant mass of the di-Higgs system
$m_{hh}$, in both the boosted and resolved categories.
%
In the case of the $p_T^{h1}$ distribution, the differences between the selection
criteria for the resolved
and boosted categories is reflected in the rightward shift of the latter.
%
The effect of PU is small in the boosted case, and generally moderate
in the resolved case, except for at large $p_T$.
%
A similar behaviour is observed in the di-Higgs invariant mass distribution.
%
Therefore we can consider the PU subtraction strategy
as validated for the purposes of this study, although
further optimisation should be possible.

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
  \includegraphics[width=0.49\textwidth]{plots/pt_H0_C2_res_comp.pdf}
  \includegraphics[width=0.49\textwidth]{plots/pt_H0_C2_bst_comp.pdf}
  \includegraphics[width=0.49\textwidth]{plots/m_HH_C2_res_comp.pdf}
  \includegraphics[width=0.49\textwidth]{plots/m_HH_C2_bst_comp.pdf}
  \caption{\small
    Comparison of the transverse momentum $p_T^h$ of the leading
    Higgs candidate (upper plots) and of the invariant mass $m_{hh}$
    of the di-Higgs system (lower plots) in the resolved
    (left) and boosted (right) categories,
    without PU and with $\la n_{PU}\ra=80$ subtracted with {\tt SoftKiller}.
}
\label{fig:mHH_PU}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
  \includegraphics[width=0.49\textwidth]{plots/D2_h0_bst_comp.pdf}
  \includegraphics[width=0.49\textwidth]{plots/tau21_h0_bst_comp.pdf}
   \caption{\small
     Comparison of the substructure variables $D_2^{(\beta)}$ (left)
     and $\tau_{21}$ (right)
     for the leading Higgs candidate in the boosted category,
   without PU and with $\la n_{\rm PU}\ra=80$ subtracted with {\tt SoftKiller}.
}
\label{fig:Substructure_PU}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

It is also interesting to quantify how
the relative differences between
signal over background distributions are modified by the inclusion of PU.
%
Considering first of all the boosted category,
in Fig.~\ref{fig:signal-vs-back-boosted} we compare
various kinematic distributions for signal and background events,
with and without PU for the leading Higgs candidate: the invariant mass, the transverse
momentum distribution $p_T$,
     the 2--to--1 subjettiness ratio $\tau_{21}$, and 
     the $k_T$ splitting scale $\sqrt{d_{12}}$, Eq.~(\ref{eq:ktsplitting}).
     %
      We verify that the relevant
      qualitative differences between signal
      and background distributions are maintained even in the presence of PU.
      %
      This is especially noticeable for the substructure variables, which
      exhibit a similar discrimination power both with and without
      PU.

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
   \includegraphics[width=0.49\textwidth]{plots/m_h0_res_comp_back.pdf}
  \includegraphics[width=0.49\textwidth]{plots/pt_h0_res_comp_back.pdf}
     \caption{\small
       Same as Fig.~\ref{fig:signal-vs-back-boosted} for the resolved category,
       this time without the jet substructure variables.
}
\label{fig:signal-vs-back-resolved}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

>>>>

In this section we describe our analysis strategy.
%
First of all we discuss the settings
for jet clustering and the strategy for jet $b$-tagging.
%
Following this we discuss the categorisation of events into different
topologies, and how the different topologies may be prioritised.
%
We motivate our choice of analysis cuts by comparing signal and background
distributions for representative kinematic variables.
%
Finally, we describe the simulation of PU and validate
the PU subtraction strategy.

@@

  The $N$-subjettiness variables $\tau_N$ are defined by clustering the constituents
  of a jet with the exclusive $k_t$ algorithm~\cite{Catani:1993hr}
  and requiring that $N$ subjets are found,
  \be
  \tau_N \equiv \frac{1}{d_0} \sum_k p_{T,k}\cdot {\rm min}\lp \delta R_{1k}, \ldots,
  \delta R_{Nk}\rp \, , \qquad d_0\equiv \sum_k p_{T,k}\cdot R \, ,
  \ee
  where $p_{T,k}$ is the $p_T$ of the constituent particle $k$ and $\delta R_{ik}$ the distance from
  subjet $i$ to constituent $k$.
  %
  In this work we use as input to the MVA the ratio of 2-subjettiness to 1-subjettiness, namely
  \be
  \label{eq:tau21}
\tau_{21} \equiv \frac{\tau_2}{\tau_1} \, ,
  \ee
  which provides good discrimination 
  between QCD jets and jets arising from the decay of
  a heavy resonance.

@@

    Large-$R$ jets are $b$-tagged by
    ghost-associating anti-$k_T$ $R=0.3$ (AKT03)
    subjets to the original large-$R$
    jets~\cite{Cacciari:2007fd,Aad:2013gja,
      ATLAS-CONF-2014-004,Aad:2015uka}.
    %
    A large-$R$ jet is considered $b$-tagged if both
    the leading and subleading AKT03 subjets, where the ordering
    is done in the subjet $p_T$, are both individually $b$-tagged,
    with the same criteria as the small-$R$ jets.
    %
     Therefore, a large-$R$ jet where the two leading
    subjets have at least one $b$-quark will be tagged
    with probability $f_b^2$.

@@

For the $b$-tagging probability $f_b$, along with
the $b$-mistag probability of light ($f_l$) and charm ($f_c$) jets,
we use the values $f_b=0.8$, $f_l=0.01$
and  $f_c=0.1$.

@@

  An event which
  contains at least two large-$R$ jets, with the two leading jets
 being $b$-tagged.
 %
 Each of these two  $b$-tagged, large-$R$ jets are 
 therefore candidates
 to contain the decay products of a Higgs boson.

@@

  An event with exactly one  $b$-tagged, large-$R$ jet, which
  is assigned to be the leading Higgs candidate.
  %
  In addition, we require at least two $b$-tagged, small-$R$ jets,
  which must be separated with respect to the large-$R$ jet
  by an angular distance of $\Delta R\ge 1.2$.
  %

@@

This is achieved as follows.
 %
First of all we perform an inclusive analysis, and optimise the
signal significance
$S/\sqrt{B}$ in each of the three categories separately, including
the MVA.
%
We find that the category with highest significance is
the boosted one,
followed by the intermediate and the resolved topologies, the latter two
with similar significance.
%
Therefore, when ascertaining which category an event is to be exclusively placed:
if the event satisfies the boosted requirements, it is assigned to
this category, else we check if it suits the intermediate
requirements.
%
It the event also fails the intermediate category
requirements, we
then check if it pass the resolved selection criteria.
%
The resulting exclusive event samples are then separately processed
through the MVA, allowing a consistent combination
of the significance of the three event categories.

@@

We now motivate the
kinematic cuts applied to the different categories, 
comparing representative kinematic distributions between
signal and background events.
%
Firstly we present results without PU, and then
discuss
the impact of PU
on the description of the kinematic
distributions.
%
In the following, all
distributions are normalized to their total integral.

In Fig.~\ref{fig:cutplots1} we show
the $p_T$ distributions
of the
  leading and subleading large-$R$ jets in the boosted category.
  %
  We observe that the background distribution
falls off more rapidly as a function of $p_T$ than the di-Higgs signal.
  %
  On the other hand, the cut in $p_T$ cannot be too strong to avoid
  a substantial degradation of signal selection efficiency,
  specially taking into account the subleading large-$R$ jet.
  %
  This comparison justifies the cut of $p_T \ge 200$ GeV
  for the large-$R$ jets that we impose in the boosted category.

@@

Another selection requirement for the boosted category is that the two
leading AKT03 subjets of the large-$R$ jet
should satisfy $p_T \ge $ 50 GeV.
%
To motivate this cut, in Fig.~\ref{fig:cutplots22}
we show the distribution in $p_T$ of the leading
and subleading AKT03 subjets in the subleading large-$R$ jet in events
corresponding to the boosted category.
%
It is clear from the comparison that the subjet $p_T$ spectrum is
relatively harder in the signal with respect to the background.
%
On the other hand, considering the subleading AKT03 subjet,
this cut in $p_T$
cannot be too harsh, to maintain a high signal selection
efficiency.
%
Therefore,
as for the previous distribution, also here 
a compromise between suppressing backgrounds but keeping a large fraction of
signal events is crucial.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
 \includegraphics[width=0.48\textwidth]{plots/pt_leadSJ_fj2_noPU.pdf}
 \includegraphics[width=0.48\textwidth]{plots/pt_subleadSJ_fj2_noPU.pdf}
 \caption{\small  Same as Fig.~\ref{fig:cutplots1} for the leading (left)
   and subleading (right) AKT03
   subjets in the subleading Higgs candidate large-$R$ jet.
}
\label{fig:cutplots22}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

Turning to the resolved category, an important aspect to account for
in the  selection
cuts is the fact that the $p_T$ distribution
of the four leading small-$R$ jets of the event can be relatively soft,
specially for the subleading jets.
%
As noted in~\cite{deLima:2014dta}, this is due to the fact
that the boost from the Higgs decay is moderate,
therefore the $p_T$ selection cuts for the small-$R$ jets cannot be too large.
%
In Fig.~\ref{fig:cutplots23}
we show the distribution in $p_T$ of the four leading
small-$R$ jets in signal and background events: we observe that both
distributions peak at $p_T \le 50$ GeV, with the signal distribution
eventually becoming dominant at large $p_T$.
%
The feasibility of triggering on four small-$R$ jets with a relatively
soft $p_T$ distribution is one of the experimental challenges for
exploiting the resolved category in this final state,
and hence the requirement that $p_T \ge 40$ GeV for
the small-$R$ jets.
%
In  Fig.~\ref{fig:cutplots23} we also show the
rapidity distribution of the the small-$R$
jets in the resolved category.
%
As expected, the production
is mostly central, more so in the case
of signal events, since backgrounds are dominated by
QCD $t$-channel exchange, therefore the
selection criteria on the jet rapidity are very efficient.

@@

The invariant mass of the di-Higgs system is another important
kinematic distribution for this process.
%
The di-Higgs invariant mass is a direct measure of the boost of the system,
which in  BSM scenarios can be substantially
enhanced, for instance due to
specific $d=6$ EFT operators~\cite{Azatov:2015oxa}.
%
One important advantage of the $b\bar{b}b\bar{b}$ final state for
di-Higgs production is that it significantly increases the reach
in $m_{hh}$ as compared to other channels with smaller branching
ratios,
such as $2b2\gamma$ or $2b2\tau$.
%
In Fig.~\ref{fig:mhh} we show the invariant mass distribution of the
reconstructed Higgs pairs,
comparing the resolved and the boosted categories.

@@

In the resolved case, we see that the distribution
in $m_{hh}$ is rather harder for the signal as compared
to the background,
and therefore one expects that cutting in $m_{hh}$ would help signal
discrimination.
%
For the boosted category the overall trend of the $m_{hh}$ distribution
is different because of the selection criteria, and the
distribution now peaks at higher values of the invariant mass.
%
In this case, signal and background distributions are not significantly
differentiated.
%
Note that at parton-level the $m_{hh}$ distribution
for signal events has a kinematic
cut-off at $m_{hh}^{\rm min}=250$ GeV, which is smeared due
to parton shower and detector resolution effects.

In Fig.~\ref{fig:pthh} we show the transverse momentum of
the di-Higgs system, $p_T^{hh}$,
for the resolved and boosted categories.
%
Once more we see that the background has a steeper fall-off in $p_T^{hh}$
than the signal, in both categories, therefore this variable
should provide additional discrimination power, motivating its inclusion
as one of the inputs for the MVA.
%
In our LO simulation the $p_T^{hh}$ distribution is generated
by the parton shower, an improved theoretical
description would require
merging higher-multiplicity
matrix elements~\cite{Maierhofer:2013sha} or matching to
the NLO calculation~\cite{Frederix:2014hta},
%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
    \includegraphics[width=0.48\textwidth]{plots/split12_h1_C1_boost.pdf} 
  \includegraphics[width=0.48\textwidth]{plots/EEC_C2_h0_C1_boost.pdf}
  \includegraphics[width=0.48\textwidth]{plots/tau21_h0_C1_boost.pdf}
  \includegraphics[width=0.48\textwidth]{plots/tau21_h1_C1_boost.pdf}
  \caption{\small Distribution of representative substructure variables
    in the boosted category at the end of the cut-based
    analysis, to be used as input to the MVA.
    %
    From top to bottom and from left to right we show  the
    $k_t$ splitting scale $\sqrt{d_{12}}$,
    the energy correlation ratio $C_2^{(\beta)}$
    and the  subjettiness ratio $\tau_{21}$ for the leading
    Higgs.
    %
    In the case of 
    $\tau_{21}$  the distributions for the subleading Higgs are also given.
 }
\label{fig:mva_substructure_1}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

From Fig.~\ref{fig:mva_substructure_1}
we observe how for these substructure variables the shapes of the signal
and background distributions reflect
the inherent differences in the internal structure of
QCD jets and jets originating from Higgs decays.
%
Signal and background distributions peak
in rather
different regions. For example, the $k_t$ splitting scale $\sqrt{d_{12}}$
peaks around 80 GeV (40 GeV) for signal (background) events, while
the distribution of the
ECF ratio $C_2^{(\beta)}$ is concentrated at small values
for signal and is much broader for background events.
%
From Fig.~\ref{fig:mva_substructure_1} we also see
the distributions of the subjettiness ratio $\tau_{12}$ are
reasonably similar
for both the leading and the subleading jets.
%

@@

Now we turn to discuss how the description of kinematic
distributions for $hh\to 4b$ signal
and background processes are
modified in the presence of pile-up.
%
To study the impact of PU,
Minimum Bias events have been generated
with {\tt Pythia8}, and then
superimposed to the signal
and background samples described in Sect.~\ref{mcgeneration}.
%
We have explored two scenarios,
one with a mean number of
PU vertices per bunch crossing of $\la n_{\rm PU}\ra=80$,
and another
with $\la n_{\rm PU}\ra=150$.
%
While the latter is closer to the HL-LHC forecasts,
we adopt $\la n_{\rm PU}\ra=80$ as our baseline,
given that it
is beyond the scope of this paper to fully optimize
the PU subtraction.

@@

In Fig.~\ref{fig:mHH_PU}
we compare the transverse momentum of the leading Higgs
candidate, $p_t^{h_1}$ and the invariant mass of the di-Higgs system
$m_{hh}$, in both the boosted and resolved categories.
%
In the case of the $p_T^{h_1}$ distribution, the differences between the selection
criteria for the resolved
and boosted categories is reflected in the rightward shift of the latter.
%
The effect of PU is small in the boosted case, and generally moderate
in the resolved case, except for at large $p_T$.
%
A similar behaviour is observed in the di-Higgs invariant mass distribution.
%
Therefore we can consider the PU subtraction strategy
as validated for the purposes of this study, although
further optimisation should still be possible.

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
  \includegraphics[width=0.49\textwidth]{plots/pt_H0_C2_res_comp.pdf}
  \includegraphics[width=0.49\textwidth]{plots/pt_H0_C2_bst_comp.pdf}
  \includegraphics[width=0.49\textwidth]{plots/m_HH_C2_res_comp.pdf}
  \includegraphics[width=0.49\textwidth]{plots/m_HH_C2_bst_comp.pdf}
  \caption{\small
    Same as Fig.~\ref{fig:m_H_PU} for
   the transverse momentum $p_T^h$ of the leading
    Higgs candidate (upper plots) and of the invariant mass $m_{hh}$
    of the di-Higgs system (lower plots) in the resolved
    (left) and boosted (right) categories.
}
\label{fig:mHH_PU}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
  \includegraphics[width=0.49\textwidth]{plots/D2_h0_bst_comp.pdf}
  \includegraphics[width=0.49\textwidth]{plots/tau21_h0_bst_comp.pdf}
  \caption{\small
    Same as Fig.~\ref{fig:m_H_PU} for the
    substructure variables $D_2^{(\beta)}$ (left)
     and $\tau_{21}$ (right)
     for the leading Higgs candidates in the boosted category.
}
\label{fig:Substructure_PU}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

It is also interesting to quantify how
the relative differences between
signal over background distributions are modified by the inclusion of PU.
%
Considering first of all the boosted category,
in Fig.~\ref{fig:signal-vs-back-boosted} we compare
various kinematic distributions for signal and background events,
with and without PU for the leading Higgs candidate: the invariant mass, the transverse
momentum distribution $p_T$,
     the 2--to--1 subjettiness ratio $\tau_{21}$, and 
     the $k_T$ splitting scale $\sqrt{d_{12}}$.
     %
      We verify that the relevant
      qualitative differences between signal
      and background distributions are maintained  in the presence of PU.
      %
      This is especially noticeable for the substructure variables, which
      exhibit a similar discrimination power both with and without
      PU.

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
      \includegraphics[width=0.49\textwidth]{plots/pt_h0_res_comp_back.pdf}
   \includegraphics[width=0.49\textwidth]{plots/m_h0_res_comp_back.pdf}
     \caption{\small
       Same as Fig.~\ref{fig:signal-vs-back-boosted} for the resolved category.
}
\label{fig:signal-vs-back-resolved}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%
